---
title: "<center> Does Your View on Disciplining Children Make You a Potential Murderer? </center>"
output: 
  html_document:
    theme: simplex
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
library(mosaic)
library(tidyverse)
library(pander)
library(DT)
library(ggrepel)
library(plotly)
library(dplyr)
library(ggplot2)
library(maps)
library(tmap)
library(leaflet)
library(htmltools)
library(car)
library(mosaicData)
library(ResourceSelection)
library(reshape2)
library(RColorBrewer)
library(scatterplot3d)
library(readr)
library(prettydoc)
library(knitr)
library(kableExtra)
library(formattable)
library(haven)

```

<!-- Directions

Come up with a chi squared test analysis of your own creation. 

* Write the Background and clearly state your question. Then collect data and perform the chi-squared test analysis in order to answer your question. Use the proper order of first, ask the question, then second, figure out the answer.

* You can use data from 
    - An R data set
    - the internet (like weather data, Facebook data, or sports data) 
    - your own experiment 
    - your own observational study
* However,     
    - Do not violate the "Human Subjects Rules" of the university: http://www.byui.edu/executive-strategy-and-planning/institutional-research/research-and-assessments/research-and-surveys

* You need a sample size large enough to meet the requirements of the chi squared test. Unfortunately, the sample size is completely dependent on the situation you select. A good rule of thumb is 10 observations per "combination-group" in your data.

-->


## Background

<br>

The [GSS Data Explorer Page](https://gssdataexplorer.norc.org/variables/vfilter) is a publicly available national source displaying the surveyed results of American Society[^1]. Within this data, I was interested in the `SPANKING` and `POLMURDR` columns. The `SPANKING` column questioned individuals if they favored spanking to discipline their child, the options being : Strongly Agree (1), Agree (2), Disagree (3), Strongly Disagree (4). The `POLMURDR` column questioned individuals if they have ever been questioned as a potential murder suspect, their options being: Yes(1) or No(2). 

In this study, we will be using the previous columns provided by GSS Data Explorer to answer the question: ***Does your views on disciplining children make you a potential murderer?***

To determine this, we will be using a **Chi-Squared Test** to determine if the preference to discipline your child is *associated or not associated* with being suspected as a potential murder. The data for this study is shown below:


```{r message=FALSE, warning=FALSE}
gss2021 <- read_csv("C:/Users/paige/OneDrive/Documents/Fall Semester 2024/MATH 325/Statistics-Notebook-master/Data/gss2021.csv")

childmurdr <- table(gss2021$SPANKING,gss2021$POLMURDR)

rownames(childmurdr) <- c("Strongly Agree", "Agree", "Disagree", "Strongly Disagree") #these replace the 1, 2, and 3

colnames(childmurdr) <- c("Yes", "No") #these replace the 1 and 2

chillM <- gss2021 %>%
  select(SPANKING, POLMURDR)

datatable(chillM, options = list(c(3,10,30)))
```



<br>

---

## Hypothesis

<br>

Aligned with our question, our null and alternative hypotheses will be stated in the following way:

$$H_0 : \text{A person's views on spanking as discipline and their status as a murder suspect are independent.} $$
$$H_a : \text{A person's views on spanking as discipline and their status as a murder suspect are associated (not independent.} $$
<br>

Additionally, we will be using this level of significance:

$$\alpha = 0.05$$

<br>

---

## Analysis

<br>

To discover the answer to our question, we will now look through graphics, tables, and conduct a chi-squared test.

<br>

---

### Graphics

To better visualize our table, we will be using a bar plot. The bar plot at a glance shows that there are significantly more non-murder suspects across all disciplinary views. However, among murder suspects, there was a higher proportion who agreed or strongly agreed with spanking compared to non-suspects with similar views. We will examine these patterns further through chi-squared testing. 

<center>

```{r}
barplot(childmurdr, beside = TRUE, legend.text= TRUE, args.legend = list(title="Should we discipline our children?", x="topright", cex=0.6, bty='n'), col = c("blue", "lightblue", "pink", "red"), xlim= c(0, 18), xlab="Have you been suspected of murder?", main = "Discplining Parents to Potential Murders of 2021")
```

```{r}
childmurdr %>%
  pander(caption = "2021 Potential Murderers", split.table=Inf)
```


</center>

<br>

---

### $\chi^2$ Test

<br>

Below are the results of our $X^2$ test.

```{r}
chi.murdr <- chisq.test(childmurdr)

pander(chi.murdr)
```

Based on the p-value of $3.37e^{-6}$, we can **reject** our null hypothesis and conclude that a person's views on spanking as discipline and their status as a murder suspect are **associated (not independent)**. This can be represented by the following: 

$$\text{P-value} = 3.37 e^{-6} < \alpha$$

<br>

---

### $\chi^2$ Requirements

<br>

In order to confirm our findings, we will take a look the requirements by analyzing the **Expected Counts** and the **Residuals**. 

<br>

There are two things to check when we look at our counts: 

1. All expected counts must be **greater than 5**

2. **How far the the residuals** happen when comparing the expected and observed counts

<br>

*Click the tabs below to see our results*

### { .tabset .tabset-pills}

#### Hide Requirements

#### Expected Counts

<br>

As we can see below, **all expected counts are greater than 5**, thus our requirements for this data is met. 

```{r}
pander(chi.murdr$expected)
```

<br>

---
  
#### Residuals

<br>

The residuals show how much our observed counts deviated from the expected counts under our null hypothesis. The largest difference between observed and expected values occurred where **individuals strongly agreed with disciplining their child and were suspected murderers**, with a residual difference of **2.828**. This relatively small difference suggests limited variability between groups. 

```{r}
pander(chi.murdr$residuals)
```

<br>

---

## Interpretation 

<br>

Based on our low p-value of $3.37e^{-6}$, we conclude that **there is an association between a person's view on spanking and their status as a murder suspect**. 

The largest raw residual appeared among individuals who **strongly agreed with spanking and were suspected murderers**, showing a deviation of about **2.83** cases above expectation.

When expressed as percentage differences between observed and expected counts:

- Strongly Agree murder suspects were about +35% higher than expected.

- Agree murder suspects were roughly +20% higher than expected.

Conversely, those who **disagreed or strongly disagreed tended to be underrepresented** among murder suspects, with deviations ranging **between -10% and -25%** compared to what independence would predict.

This percentage framing emphasizes that the observed association is not just statistically significant but also practically meaningful, as particular disciplinary views appear disproportionately among those suspected of murder.


These findings suggest that investigators might consider a suspect's disciplinary attitudes during their investigation process, particularly when determining which individuals warrant more thorough questioning.  

<br>

---

### Sources

[^1]:[About GSS Data Explorer Page](https://gssdataexplorer.norc.org/about)

- Formatting Inspiration 

  - [Decade of Movies Chi-Squared Test Example Analysis](file:///C:/Users/paige/OneDrive/Documents/Fall%20Semester%202024/MATH%20325/Statistics-Notebook-master/Analyses/Chi%20Squared%20Tests/Examples/movies.html)

