---
title: "Rent"
output: 
  html_document:
    theme: spacelab
    code_folding: hide
    toc: true
    toc_float: true
---

<!-- Welcome to the first Analysis for Math 325! 
The goal of this first assignment is to have you produce a graphical summary and a numerical summary of the "Rent" data set. The fun part is that you get to decide what type of graph and what type of numerical summary to create! You will need to read through the "green text" instructions of this document to find the places where you are supposed to enter R-code in order to complete the assignment. -->


**Week 1 Analysis**

```{r, message=FALSE, warning=FALSE}
# This gray area is called an "R-chunk". 
# These library commands install some powerful functions for your use later on.
library(mosaic)
library(pander)
library(tidyverse)
library(DT) 

# This read_csv command reads in the "Rent" data set into an object called "Rent"
Rent <- read_csv("../Data/Rent.csv")

# To load this data set into your R-Console do the following:
#  1. From your top file menu select "Session -> Set Working Directory -> To Source File Location"
#  2. Press the green "play button" in the top right corner of this gray box (which is called an "R-chunk").
#  3. Then in your "Console" window of 
```

<br />

## Background

Here is a data table showing the available approved housing apartment options at BYU-Idaho for single students. There are 122 entries comprising 57 female and 65 male apartment options.

```{r, warning=FALSE}
# Code to get you started. 
# View(...) works great in the Console, but datatable(...) must be
# used instead within an R-chunk.
datatable(Rent, options=list(lengthMenu = c(3,10,30)), extensions="Responsive")
```


## Graphic

<!-- Present an interesting graphic using the Rent data set. Do not create the same graphics that are shown in the [Good Example Analysis](./Analyses/StudentHousing.html), but you can certainly use those graphics as a guide to inspire you on what to do yourself. -->


```{r}
# Use this R-chunk to...
# ...Create an interesting graphic using the Rent data set.

ggplot(Rent, aes(x=Deposit))+geom_histogram(binwidth=15, fill="dodgerblue4", color="dodgerblue3")+ggtitle("The Number of Residents Based on Rent Deposit")+ xlab("Deposit Cost (in U.S. Dollars)")+ylab("Number of Residents")+theme(plot.title=element_text(size=20,color="dodgerblue",face="bold",family="serif"),axis.title.x = element_text(size=16), axis.title.y = element_text(size=16),axis.text.x = element_text(size=12),axis.text.y = element_text(size=12))

```


<!-- Write a few statements here that explain your graphic and what it shows. -->

The following plot shows that the majority of residents pay between $80 and $150 for their rent deposit. This is apparent because of the clustering of bars within that range of deposit cost.


Other bars in the plot that are not within this range are very short, such examples being at $50 and $300. The most frequent rent deposit cost is around $150, as shown by being the tallest bar in the plot. 




```{r}
# Use this R-chunk to...
# ...compute and display a meaningful table of numerical summaries supporting your above graphic.

pander(summary(Rent$Deposit), caption ="<span style='color:dodgerblue;font-family: serif;'><strong>Summary of Rent Deposits</strong></span>")

```

<!-- Write a few statements here that introduce your table of numerical summaries and what they show. -->

This table shows the numerical summary of the rent deposits costs. The cheapest deposit option being $50. The most expensive deposit option being $300. According to the table, the average cost for rent deposit is $124. 

## Conclusion

Based on the average rent deposit, most residents tend to favor the lower end of deposit costs. While it's uncommon for students to pay the maximum deposit of $300, it's equally rare for them to opt for the minimum $50 deposit.


The skew towards lower deposit amounts suggests that properties might lower their deposit costs within this range to cater towards more residents in the future. 



