---
title: "<center>College Student's Food Perception</center>"
output: 
  html_document:
    theme: sandstone
    code_folding: hide
---

```{r, warning=FALSE, message=FALSE}
# From your file menu (at the top of R-Studio) select:
# "Session -> Set working directory -> To source file location"
# Then play this chunk to get the data into R.
library(mosaic)
library(car)
library(DT)
library(pander)
library(readr)
library(knitr)
library(kableExtra)
library(formattable)
library(ggplot2)

food <- read_csv("../../Data/food.csv") #food.csv is in the Data folder...
```


----

<!-- Instructions:

 Use the Food dataset and a Kruskal-Wallis Test(s) to answer an interesting question(s) that you come up with.
 
 About the dataset:
 
  * GPA - numerical, self-reported current college GPA

  
  * weight - numerical, self-reported weight in lbs


  * gender - "Female", "Male"


  * breakfast - "Cereal", "Donut" 
      The participants were shown a picture of "Cereal" and a "Donut" and asked which one of these pictures they associate with the word "breakfast."


  * calories_chicken - guessing calories in chicken piadina from the options of:
     1 - 265 
     2 - 430 
     3 - 610 
     4 - 720
    (the variable shows the actual number of calories participants selected)


   * calories_day - Importance of consuming calories per day
     1 - i dont know how many calories i should consume 
     2 - it is not at all important 
     3 - it is moderately important 
     4 - it is very important


   * comfort_food - List 3-5 comfort foods that come to mind. 
     Open ended 


   * comfort_food_reasons - What are some of the reasons that make you eat comfort food?      (i.e., anger, sadness, happiness, boredom, etc) - list up to three
     Open ended


   * comfort_food_reasons_coded
     1 - stress
     2 - boredom
     3 - depression/sadness
     4 - hunger
     5 - laziness
     6 - cold weather
     7 - happiness 
     8 - watching tv
     9 - none 


    * cook - how often do you cook?
     1 - Every day 
     2 - A couple of times a week 
     3 - Whenever I can, but that is not very often  
     4 - I only help a little during holidays 
     5 - Never, I really do not know my way around a kitchen


    * cuisine - what type of cuisine did you eat growing up?
     1 - American
     2 - Mexican.Spanish
     3 - Korean/Asian
     4 - Indian
     5 - American inspired international dishes
     6 - other


    * diet_current - describe your current diet
     open ended

	
    * diet_current_coded
     1 - healthy/balanced/moderated/
     2 - unhealthy/cheap/too much/random/
     3 - the same thing over and over
     4 - unclear


    * drink - which picture do you associate with the word “drink”?
     1 - orange juice
     2 - soda 


    * eating_changes  - Describe your eating changes since the moment you got into college?
     Open ended 


    * eating_changes_coded
     1 - worse
     2 - better
     3 - the same
     4 - unclear


    * eating_changes_coded1
     1 - eat faster
     2 - bigger quantity
     3 - worse quality 
     4 - same food
     5 - healthier
     6 - unclear
     7 - drink coffee 
     8 - less food
     9 - more sweets
     10 - timing 
     11 - more carbs or snacking
     12 - drink more water
     13 - more variety


    * eating_out - frequency of eating out in a typical week 
     1 - Never 
     2 - 1-2 times 
     3 - 2-3 times 
     4 - 3-5 times 
     5 - every day


    * employment - do you work? 
     1 - yes full time 
     2 - yes part time 
     3 - no
     4  - other


    * ethnic_food - How likely to eat ethnic food 
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 


    * exercise - how often do you exercise in a regular week?
     1 - Everyday 
     2 - Twice or three times per week 
     3 - Once a week
     4 - Sometimes 
     5 - Never


    * father_education - 
     1 - less than high school 
     2 - high school degree 
     3 - some college degree 
     4 - college degree 
     5 - graduate degree 


    * father_profession - what is your father profession?
     Open ended


    * fav_cuisine - What is your favorite cuisine?
     Open ended


    * fav_cuisine_coded
     0-none
     1 - Italian/French/greek
     2 - Spanish/mexican
     3 - Arabic/Turkish
     4 - asian/chineses/thai/nepal
     5 - American
     6 - African 
     7 - Jamaican
     8 - indian


    * fav_food - was your favorite food cooked at home or store bought? 
     1 - cooked at home 
     2 - store bought 
     3 - both bought at store and cooked at home


    * food_childhood - what was your favorite childhood food?
     Open ended



    * fries - which of these pictures you associate with word fries? 
     1 - Mcdonald’s fries
     2 - home fries

 
    * fruit_day - How likely to eat fruit in a regular day 
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 


    * grade_level – 
     1 - freshman 
     2 -Sophomore 
     3 - Junior 
     4 - Senior
     
      
    * greek_food - How likely to eat greek food when available?
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 
     
      
    * healthy_feel – how likely are you to agree with the following statement: “I feel very healthy!” ?
      1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
      
      
    * healthy_meal – what is a healthy meal? Describe in 2-3 sentences.
      Open ended
      
      
    * ideal_diet – describe your ideal diet in 2-3 sentences
      Open ended
      
      
    * Ideal_diet_coded
      1 – portion control
      2 – adding veggies/eating healthier food/adding fruit	
      3 – balance
      4 – less sugar
      5 – home cooked/organic
      6 – current diet
      7 – more protein
      8 – unclear
      
      
     * income
      1 - less than $15,000 
      2 - $15,001 to $30,000 
      3 - $30,001 to $50,000 
      4 - $50,001 to $70,000 
      5 - $70,001 to $100,000 
      6 - higher than $100,000
      
     * indian_food – how likely are you to eat indian food when available
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * Italian_food – how likely are you to eat Italian food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * life_rewarding – how likely are you to agree with the following statement: “I feel life is very rewarding!” ?
      1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
      
      
     * marital_status
      1 -Single 
      2 - In a relationship 
      3 - Cohabiting 
      4 - Married 
      5 - Divorced 
      6 - Widowed
      
      
     * meals_dinner_friend – What would you serve to a friend for dinner?
      Open ended
      
      
     * mothers_education 
      1 - less than high school 
      2 - high school degree 
      3 - some college degree 
      4 - college degree 
      5 - graduate degree
      
      
     * mothers_profession – what is your mother’s profession? 
      
      
     * nutritional_check - checking nutritional values frequency 
      1 - never 
      2 - on certain products only 
      3 - very rarely 
      4 - on most products 
      5 - on everything
      
      
     * on_off_campus – living situation
      1 - On campus 
      2 - Rent out of campus 
      3 - Live with my parents and commute 
      4 - Own my own house
      
      
     * parents_cook - Approximately how many days a week did your parents cook? 
      1 - Almost everyday 
      2 - 2-3 times a week 
      3 - 1-2 times a week 
      4 - on holidays only 
      5 - never
      
      
     * pay_meal_out - How much would you pay for meal out? 
      1 - up to $5.00 
      2 - $5.01 to $10.00 
      3 - $10.01 to $20.00 
      4 - $20.01 to $30.00 
      5 - $30.01 to $40.00 
      6 - more than $40.01 
      
      
     * Persian_food - How likely to eat persian food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * self_perception_weight - self perception of weight 
      6 - i dont think myself in these terms 
      5 - overweight 
      4 - slightly overweight 
      3 - just right 
      2 - very fit 
      1 - slim 
      
      
     * Which of the two pictures you associate with the word soup?
      1 – veggie soup
      2 – creamy soup
      
      
     * sports - sports – do you do any sporting activity?
      1 - Yes 
      2 - No 
      99 – no answer
      
      
     * thai_food - How likely to eat thai food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * tortilla_calories - guessing calories in a burrito sandwhich from Chipolte?
      1 - 580 
      2 - 725 
      3 - 940 
      4 - 1165
      
     * turkey_calories - Can you guess how many calories are in the foods shown below?
     (Panera Bread Roasted Turkey and Avocado BLT)
      1 - 345 
      2 - 500 
      3 - 690 
      4 - 850
      
      
     * type_sports – what type of sports are you involved?
      Open-ended
      
     * veggies_day - How likely to eat veggies in a day? 
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4- likely 
      5 - very likely
      
     * vitamins – do you take any supplements or vitamins?
      1 – yes
      2 – no
      
     * waffle_calories - guessing calories in waffle potato sandwhich 
      1 - 575 
      2 - 760 
      3 - 900 
      4 - 1315


Note this dataset and description come from: © 2020 Kaggle Inc for further details visit:

[Food choices on Kaggle](https://www.kaggle.com/borapajo/food-choices?select=food_coded.csv)


If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->




<!--End your comments to your critiquers before this line.-->



<!--Begin writing your Analysis after this point.-->

## Background

## {.tabset .tabset-pills}

The following experiment conducted consisted of a survey that went around a particular college campus asking students about their GPA and Weight as well as a variety of questions regarding food. There were questions asking about their comfort food, the likely hood of eating a specific cuisine, and if a health option or a junk food option came to mind when asking about a specific food. 

In this study, we will be specifically examining the student's weight as well as their perception of food to answer the following question:

***Does the perception of food among students at this college result in a change in their weight?***

<div style="font-size:.8em;color:#888888;">
Alternatively: If a student thinks healthier, do they have a lower weight? Or, if a student thinks unhealthy do they have a higher weight?
</div>

<br><br>

*For the data table of the following experiment and more information on the data table, click `Show Data Table`:*

<br>

### Hide Data Table

<br>

---

### Show Data Table 

<div>


The survey showed individuals two pictures to choose from when it came to associating foods with a specific words. The table below shows the different questions the survey asked in regards to word food association: 

<center>

```{r message=FALSE, warning=FALSE}
questions <- c("Which one of these pictures do you associate the word 'breakfast'?",
               "Which picture do you associate with the word 'drink'?",
               "Which of these pictures you associate with word 'fries'?",
               "Which of the two pictures you associate with the word 'soup'?")

foodoptions <- c("Cereal /  Donut",
                 "1- Orange Juice / 2- Soda",
                 "1- McDonald's Fries / 2- Home Fries",
                 "1- Veggie Soup / 2- Creamy Soup")

collegestusurvey <- data.frame( Question = questions, Options = foodoptions, stringsAsFactors = FALSE)

kable(collegestusurvey, col.names = c("Questions", "Food Options"), caption ="Food Survey")
```

    
</center>


<br>

Thus, their answer was either expressed as a number or the specific word, one option being healthier than the other. If they chose the junk food option out of the two, they are rewarded a point that would then be counted for in their **Food Perception Score**. This would then result in scores ranging from 0-4, ***0 being the healthy thinking and 4 being junk food thinking***. The **Food Perception Score** column is highlighted in green in the table to display each student's score. 


     


```{r}
Foood <- food %>%
  select(weight, breakfast, fries, drink, soup) %>%
  filter(!is.na(suppressWarnings(as.numeric(weight))) + 
           is.na(drink)) %>%
  mutate(`Food Perception Score` = ifelse(breakfast == "Cereal", 0, 
                               ifelse(breakfast == "Donut", 1, 0)) +
                        ifelse(fries == "1", 1, ifelse(fries == "2", 0, 0)) + 
                        ifelse(drink == "orange juice", 0, ifelse(drink == "soda", 1, 0)) +
                        ifelse(soup == "1", 0, ifelse(soup == "2", 1, 0))) %>%
  mutate(weight = as.numeric(weight))

datatable(Foood, options=list(pageLength = 10)) %>%
  formatStyle('Food Perception Score', backgroundColor = 'lightgreen')

```

<br>

</div>

---

<br>

## Hypothesis 

To find the answers for our question above, we will be using a **Kruskal-Wallis Rank Sum Test**. Thus, we will use the following hypotheses: 

<br>

<center>

$$H_0 : \text{All samples are from the same distribution.}$$

$$H_a : \text{At least one sample's distribution is stochastically different.}$$

<br>

In more depth, the null hypothesis states that the weights of college students, regardless of their food perception (healthy or unhealthy), **come from the same distribution at this specific college**. While the alternative hypothesis suggests that **at least one** type of food perception (healthy or unhealthy) results in a **different weight distribution among college students at this particular school**.

<br>

Additionally, our level of significance will be: 




$$ \alpha = 0.05 $$

</center>


---

## Analysis

## {.tabset .tabset-fade}

### Graphical/ Numerical Summaries

The following box plot helps to show us which distribution is the **highest** out of all the **Food Perception Scores**. The different shades of green show us the range in **Food Perception Scores**, *0 being the healthy thinking and 4 being junk food thinking*. 

According the box plot, ***college students with a Food Perception Score of 3 have the highest distribution in our sample***. Additionally, the five number summary of each box plot is given below the box plot. 

<center>

```{r}
boxplot(weight ~ `Food Perception Score`, data=Foood, 
        col=c("darkseagreen1","palegreen","palegreen3","palegreen4","mediumseagreen"), xlab="Food Perception Score", main="Weight of Surveyed College Students", ylab="Weight (lbs)")
stripchart(weight ~ `Food Perception Score`, data=Foood, pch=16,vertical=TRUE, add=TRUE, col="gray28")
```


</center>

```{r}
Foodie <- favstats(weight ~ `Food Perception Score`, data = Foood)[,-10]

Foodie2 <- as.data.frame(Foodie)

kable(Foodie2)
```


---


### Kruskal-Wallis Rank Sum Test

According to our Kruskal-Wallis test shown below, we have a p-value of **0.06521**.


```{r message=FALSE, warning=FALSE}
kruskal.test(weight ~ `Food Perception Score`, data=Foood) %>%
  pander()
```


$$ p-value = 0.06521 > \alpha $$

Due to our high p-value, we are able to conclude that we have **insufficient** evidence to reject the null hypothesis. 

<br>

---

## Interpretation 

Even though the Kruskal-Wallis test shows us one thing, the visuals provided by the box plot tell us another. Student's with a Food Perception Score of 3 had the **highest distribution** as well as the **highest median of 175** in our sample. Therefore, appearing to show us that there is at least one distribution that differs from the rest as stated in our alternative hypothesis.  

However, with a p-value of **0.06521** slightly above our significance level (0.05), we have **insufficient evidence** to reject the null hypothesis for the surveyed school's student population. This suggests that the way a student perceives food may not significantly affect the student's weight distribution at this particular institution.

---

<u> Sources: </u>

- ChatGPT for helping me to figure out how to put in cool stylistic things such as the colored column in my data table

- Formatting Inspiration

    - [Student Wages - Kruskal-Wallis Example](file:///C:/Users/paige/OneDrive/Documents/Fall%20Semester%202024/MATH%20325/Statistics-Notebook-master/Analyses/Kruskal-Wallis%20Test/Examples/wagesKruskal.html)
    
    - [Pig Birth Weights - Kruskal-Wallis Example](file:///C:/Users/paige/OneDrive/Documents/Fall%20Semester%202024/MATH%20325/Statistics-Notebook-master/Analyses/Kruskal-Wallis%20Test/Examples/pigweightsKruskal.html#)







